In order to use the functions contained within this package the following R packages must be installed in the client side:
  
```{r install_all, eval=FALSE}
install.packages("DSIOpal")
install.packages('dsBaseClient', repos=c(getOption('repos'), 
                                         'http://cran.obiba.org'), dependencies=TRUE)
devtools::install_github("isglobal-brge/dsOmicsClient", dependencies = TRUE)
```

The package dependencies are then loaded as follows:
  
```{r requiredRPackages}
library(DSOpal)
library(dsBaseClient)
library(dsOmicsClient)
```

**Notes**:
  
  For advanced users willing to use `DSLite`, the server side packages needs to be installed as well:
  
```{r install_resourcer, eval=FALSE}
install.packages(c("resourcer", "DSLite"), dependencies = TRUE)
install.packages("dsBase", repos = c("https://cloud.r-project.org", 
                                     "https://cran.obiba.org"), dependencies = TRUE)
```

We refer to [this chapter](https://isglobal-brge.github.io/resource_bookdown/dslite-datashield-implementation-on-local-datasets.html) of our bookdown to a more detail description about how to work with DataSHIELD in a serverless environment.