The same analysis can be performed for all features (e.g. CpGs) just avoiding the `feature` argument. This process can be parallelized using `mclapply` function from the `multicore` package.


```{r multiple_cpg, eval=FALSE}
ans <- ds.lmFeature(model = ~ diagnosis + Sex, 
                    Set = "methy",
                    datasources = conns,
                    mc.cores = 20)
```


This method corresponds to the **pooled analysis** approach and can be very time consiming since the function repeatedly calls the DataSHIELD function `ds.glm()`. We can adopt another strategy that is to run a glm of each feature independently at each study using `limma` package (which is really fast) and then combine the results (i.e. **meta-analysis** approach). 


```{r limma_methy}
ans.limma <- ds.limma(model = ~ diagnosis + Sex,
                      Set = "methy", 
                      datasources = conns)
```

Then, we can visualize the top genes at each study (i.e server) by 

```{r show_limma_methy}
lapply(ans.limma, head)
```

The annotation can be added by using the argument `annotCols`. It should be a vector with the columns of the annotation available in the `ExpressionSet` or `RangedSummarizedExperiment` that want to be showed. The columns of the annotation can be obtained by

```{r show_annot_cols}
ds.fvarLabels("methy")
```

Then we can run the analysis and obtain the output with the chromosome and gene symbol by:
  
  
```{r remove_ans_limma, eval=FALSE, echo=FALSE}
ds.rm("ans.limma")
```



```{r limma_methy_annot}
ans.limma.annot <- ds.limma(model = ~ diagnosis + Sex,
                            Set = "methy", 
                            annotCols = c("CHR", "UCSC_RefGene_Name"),
                            datasources = conns)
```

```{r show_limma_methy_annot}
lapply(ans.limma.annot, head)
```


Then, the last step is to meta-analyze the results. Different methods can be used to this end. We have implemented a method that meta-analyze the p-pvalues of each study as follows:
  
```{r meta_p}
ans.meta <- metaPvalues(ans.limma)
ans.meta
``` 

We can verify that the results are pretty similar to those obtained using pooled analyses. Here we compute the association for two of the top-CpGs:
  
```{r one_cpg_val}
res1 <- ds.lmFeature(feature = "cg13138089", 
                     model = ~ diagnosis + Sex, 
                     Set = "methy",
                     datasources = conns)
res1

res2 <- ds.lmFeature(feature = "cg13772815", 
                     model = ~ diagnosis + Sex, 
                     Set = "methy",
                     datasources = conns)
res2
```


We can create a QQ-plot by using the function `qqplot` available in our package.

```{r qqplot}
qqplot(ans.meta$p.meta)
```

Here In some cases inflation can be observed, so that, correction for cell-type or surrogate variables must be performed. We describe how we can do that in the next two sections.
