---
title: "Non-disclosive federated omic data analysis of EGA hosted RNAseq data"
author:
- name: Xavier Escribà Montagut
  affiliation:
  - &isglobal Bioinformatics Research Group in Epidemiolgy (BRGE), Barcelona Insitute for Global Health (ISGlobal)
  email: xavier.escriba@isglobal.org
- name: Juan R. Gonzalez
  affiliation:
  - &isglobal Bioinformatics Research Group in Epidemiolgy (BRGE), Barcelona Insitute for Global Health (ISGlobal)
  - &uab Department of Mathematics, Autonomous University of Barcelona (UAB)
  email: juanr.gonzalez@isglobal.org
date: "`r Sys.Date()`"
package: "`r pkg_ver('dsOmics')`"
output:
  BiocStyle::html_document:
    number_sections: true
    toc: yes
    fig_caption: yes
    toc_float: true
bibliography: ega_Vignette.bib
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Connect to server

```{r}
require('DSI')
require('DSOpal')
require('dsBaseClient')

builder <- DSI::newDSLoginBuilder()
builder$append(server = "study1",
   url = "https://opal-demo.obiba.org/",
   user = "dsuser", password = "password",
   driver = "OpalDriver")
logindata <- builder$build()
connections <- DSI::datashield.login(logins = logindata)
```

# Assign resources and resolve them

```{r}
# Assign resources to the R session
datashield.assign.resource(connections, "RNAseq_resource", "RSRC.RNAseq_EGA")
datashield.assign.resource(connections, "RNAseq_pheno_resource", "RSRC.RNAseq_pheno_EGA")
# Resolve resources
datashield.assign.expr(conns = conns, symbol = "RNAseq", expr = quote(as.resource.object(RNAseq_resource)))
datashield.assign.expr(conns = conns, symbol = "RNAseq_pheno", expr = quote(as.resource.data.frame(RNAseq_pheno_resource)))
```

#  Create RangedSummarizedExperiment (RSE) with the two tables

In order to create a RSE, the RNAseq file, has to comply with the following:

+ First column: Entrez identificator
+ Second column: Annotated symbol
+ All the remaining columns: One col. per individual

```{r}
ds.colnames('RNAseq_pheno')
```

The selected resource complies with the requirements. We can now proceed to merge the RNAseq with the phenotypes file into a RSE.

```{r}
ds.createRSE('RNAseq', 'RNAseq_pheno')
```

# Study of RSE

```{r}
ds.dim('RSE')
name.features <- ds.featureNames("RSE")
lapply(name.features, head)
name.vars <- ds.featureData("RSE")
lapply(name.vars, head, n=15)
ans.gender <- ds.limma(model =  ~ gender, 
                       Set = "RSE", type.data = "RNAseq", 
                       sva = FALSE)
```

